On Devon Allen's Disqualification at the 2022 Track and Field World Championship

### 2023-04-21 Manuscript/manuscript.tex
clusrank ds

\subsection{Beyond World Championship}
This section was motivated from a qualitative analysis that started by looking
at how United States athletes reacted at the 2022 United States Track and Field
Championships. From June 23-26 2022, USA held its Track
and Field Championship to decide who to send to the World Championships at Hayward 
Field, the exact same venue as used for the World Track and Field Championship a month 
later.  Thus, a baseline is able to be established for the four USA athletes who 
competed in at least one round of both events: Trey Cunningham, Daniel Roberts, 
Grant Holloway, and Devon Allen. Every athlete reacted faster in all the World 
Track and Field Championship races compared to the USA Track and Field 
Championship races. Devon Allen raced three times at the USA Track and
Field Championship and three times at the World Track and Field  Championship. 
His reaction times were significantly faster in the World Track and
Field Championship compared to the USA Track and Field Championship as were his
teammates.  This comparison suggests that the timing system used between the two
meets may have been meaningfully different to the point where Devon Allen was
disqualified more because of faulty equipment rather than because he reacted
\jy{cross-reference to the table by its label}
too quickly.  Table~\ref{fig:USAvsWorld} below highlights the difference in reaction times for
Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. ``USA" denotes
the USA Track and Field Championships, ``World" denotes the World Track and Field
Championships,   All numbers listed in the table are the reaction
time for the respective athletes in seconds.

@@

@@

\jy{Undefined notations}
The test statistic of the DS method of the Wilcoxon rank-sum statistic is:
$Z = \frac{S - E(S)}{\sqrt{\hat{\text{Var}}(S)}}$ where 
$W^* = \frac{1}{N} + \sum_{i=1}^{N} \delta_{i}^{*} R_{i}^{*}$, $R_{i}^{*}$ is
the rank of $X_{i}^{*}$. S is given as $S = \mathbb{E}[W^* | X, \delta]$ and
$E(S) = E(W^*) = \frac{1}{2} \sum_{i=1}^{N} \frac{n_i}{N}$.
The asymptomatic distribution of the Wilcoxon rank-sum statistic was found by
\citet{datta2005rank} and the R package ``clusrank" allows for its implementation.
\citep{jiang2017wilcoxon}.
---Please see manuscript around line 464 for comments regarding this section---

%I included additional statistics from the paper, I am unsure of how in depth
%This explanation needs to be

%$F_{i}(x) = \frac{1}{n_i} \sum_{k=1}^{n_i} \mathbb{I}(X_{i,k} \leq x)$

%$F_{i}(x-) = \frac{1}{n_i} \sum_{k=1}^{n_i} \mathbb{I}(X_{i,k} < x)$

%$S = \frac{1}{N + 1} \sum_{i=1}^{N}  \sum_{k=1}^{n_i} \frac{\delta_{ik}}{n_i} \left[ 1 + \frac{1}{2} \sum_{j \neq i} \frac{1}{n_j}
%\left( F_j(X_{ik}) + F_j(X_{ik-}) \right) \right]$

Using the \texttt{clusWilcox.test()} function, a Wilcoxon rank sum
test was performed with the cluster set to be the Athlete and group set to Year
using the DS method. The DS method is preferable when the cluster sizes are
informative \citep{jiang2017wilcoxon}.  In this case, athletes who have more
data are generally thought to be faster, as they have made it further in the
competition.  Clusters range from two to six, with two indicating that at each
competition they only made it to the heats, but a cluster size of six indicates
that they made the finals in both competitions.  While the number of times
athletes compete is based off their overall time not their reaction time, the
two are correlated \citep{delalija2008reaction}.

@@

\bibliographystyle{chicago}
\bibliography{citations.bib}

>>>>

\subsection{Beyond World Championship}
This section was motivated from a qualitative analysis that started by looking
at how United States athletes reacted at the 2022 United States Track and Field
Championships. From June 23-26 2022, USA held its Track
and Field Championship to decide who to send to the World Championships at Hayward 
Field, the exact same venue as used for the World Track and Field Championship a month 
later.  Thus, a baseline is able to be established for the four USA athletes who 
competed in at least one round of both events: Trey Cunningham, Daniel Roberts, 
Grant Holloway, and Devon Allen. Every athlete reacted faster in all the World 
Track and Field Championship races compared to the USA Track and Field 
Championship races. Devon Allen raced three times at the USA Track and
Field Championship and three times at the World Track and Field  Championship. 
His reaction times were significantly faster in the World Track and
Field Championship compared to the USA Track and Field Championship as were his
teammates.  This comparison suggests that the timing system used between the two
meets may have been meaningfully different to the point where Devon Allen was
disqualified more because of faulty equipment rather than because he reacted
\jy{cross-reference to the table by its label}
too quickly.  Table~\ref{fig:USAvsWorld} highlights the difference in reaction times for
Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. ``USA" denotes
the USA Track and Field Championships, ``World" denotes the World Track and Field
Championships,   All numbers listed in the table are the reaction
time for the respective athletes in seconds.

@@

\jy{You have two such datasets in the analysis. Use one paragraph for each.}

@@

To test the conjecture that the 2022 World Champion timing device may have led
to faster recorded reaction times, we compare the reaction times of the same
athletes who have attended both the 2022 World Champion and other competitions.
In this setting, we have clustered data with subunit grouping. In particular,
each athlete is a cluster and the multiple reaction times from the same athlete
can be from either the 2022 World Championship or otherwise.
Let $X_{ij}$ be the $j$th reaction time of athelete~$i$, $i = 1, \ldots, n$,
$j = 1, \ldots, m_i$ where $m_i$ is the number of observations from
athelete~$i$. Let $\delta_{ij}$ be the group indicator of $X_{ij}$; $\delta_{ij}
= 1$ if $X_{IJ}$ is in group~1 and $\delta_{ij} = 0$ otherwise. Atheletes are
assumed to be independent, while subunit observations from the same athelete are
not. The null hypothesis $H_0$ to be tested is that there is no difference
between the two groups; i.e., the distribution of $X_{ij}$ remains the same
regardless of the group indicator $\delta_{ij}$.

\citet{datta2005rank} proposed an extension of the Wilcoxon rank-sum test to
clustered data with subnit-level grouping. The test is designed based on a
within-cluster resampling principle. Consider randomly picking one observation
from each cluster to form a pseudo-sample. Let $X_i^*$ be a random pick from the
$i$th cluster in the pseudo-sample and $\delta_i^*$ its group indicator. The
Wilcoxon rank-sum statistic for the pseudo-sample is
\[
W^* = \frac{1}{n + 1} + \sum_{i=1}^{n} \delta_{i}^{*} R_{i}^{*},
\]
where $R_{i}^{*}$ is the rank of $X_{i}^{*}$ in the psudo-sample.
The test statistic $S$ is the average of $W^*$ averaged over all possible
pseudo-samples conditioning on the observed data and group indicators.
The mean and variance of $S$ under $H_0$ can be derived so that $S$ can be
standardized to form a statistic $Z$ which follows standard normal distribution
asymptotically \citep[p.910]{datta2005rank}. An implementation of this method is
available from the \texttt{clusWilcox.test()} function with
\texttt{method = 'ds'} from R package \texttt{clusrank}
\citep{jiang2017wilcoxon}.
%% In this case, athletes who have more
%% data are generally thought to be faster, as they have made it further in the
%% competition.  Clusters range from two to six, with two indicating that at each
%% competition they only made it to the heats, but a cluster size of six indicates
%% that they made the finals in both competitions.  While the number of times
%% athletes compete is based off their overall time not their reaction time, the
%% two are correlated \citep{delalija2008reaction}.

@@

\jy{Go over the references to see clues to clean in the bib source}
\bibliographystyle{chicago}
\bibliography{citations.bib}
